Online learning methods for financial time series forecasting

### 2017-11-19 slides/defense-slides.tex
SATURDAY

The relationship between non-stationary time series can be modelled if exists a linear combination of them that turns out to be stationary.
%Financial time series are mainly non-stationary.  However, one or more linear
%combinations of non-stationary time series can be stationary even though
%individually they are not. 
When this happens it is said they have a long-run
equilibrium relationship and the variables are {\color{red} cointegrated}.
\
% The Vector Error correction model (VECM) is used to model cointegrated time series but is used only with batch data and rarely used with high frequency data.

@@

%\subsection{Proposal}
%\frametitle{Proposal}
%\begin{block}{}
%To use VECM with high frequency data where the best parameters are updated when new data is available and do this is done in a fast way.
%\section{Background}
%%\frametitle{Integration }
%%\begin{block}{Integration of order $d$}
%%A time series $\mathbf{y}$ is said to be integrated of order $d$ if after
%%differentiating the variable $d$ times, we get an stationary process, more precisely:
%%\[
%%(1-L)^d \mathbf{y} \sim \text{I(0)} \, ,
%%\]
%%\noindent where I(0) is a stationary time series and $L$ is the lag operator:
%%\[
%%(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
%%\]
%%\subsection{Integration example}
%%\frametitle{Financial Time Series}
%%\begin{columns}
%%\column[t]{5cm}
%%High frequency financial time series are commonly an I(1) process 
%%\begin{center}
%%\includegraphics[width=0.85\textwidth]{img/EURUSD}
%%\end{center}
%%\column[t]{5cm}
%%Since their differences are stationary (I(0) process).
%%\begin{center}
%%\includegraphics[width=0.85\textwidth]{img/DEURUSD}
%%\end{center}
%%\end{columns}
%\subsection{Cointegration}
%\frametitle{Cointegration}
%\begin{block}{Cointegration definition}
%Let $\mathbf{y} = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$ be a set of $l$
%time series of order I(1) which are said to be cointegrated if a vector
%$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the
%time series,
% \mathbf{Z}_t:= \beta^\top \mathbf{y} = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
% \text{I(0)}\, .
%In other words, a set of I(1) variables is said to be cointegrated if
%a linear combination of them exists which is I(0).
%\subsection{Vector Autorregresive models}
%\frametitle{VECM: Vector Error Correction model}
%VECM is a special form of a Vector Autorregresive model (VAR) for I(1)
%variables $\mathbf{y}$ that are also cointegrated. Let 
%\label{eq:variables}
%\mathbf{y}_t = 
%\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t}
%\end{bmatrix}^\top \, \qquad t=p+1, \dots N. 
%\begin{block}{VAR}
%VAR($p$) model describes the behaviour of a set of $l$
%endogenous variables as a linear combination of their last $p$ values:
%\label{eq:var}
% \mathbf{y}_t = \phi_1 \mathbf{y}_{t-1}  + \dots +   \phi_p\mathbf{y}_{t-p}
%+ \mathbf{c} + \mathbf{\epsilon}_t \, ,
%\noindent where  ${\phi_1,\dots,\phi_p}$ are $l \times l$
%matrices of real coefficients,
%$\mathbf{\epsilon}_{p+1},\dots,\mathbf{\epsilon}_N$ are error terms,
%$\mathbf{c}$ is a constant vector and $N$ is the total number of samples.
%\frametitle{VECM: Vector Error Correction model}
%\begin{block}{VECM}
%It is obtained by replacing the form {\color{red} $\Delta
%\mathbf{y}_t = \mathbf{y}_t - \mathbf{y}_{t-1}$} in the VAR form.
% \label{eq:vec}
% \Delta \mathbf{y}_t = 
% \underbrace{ \Omega\mathbf{y}_{t-1}}_\text{Error correction term} + 
% \sum_{i=1}^{p-1}
%\phi_i^* \Delta \mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
%\end{equation}}

>>>>

The relationship between non-stationary time series can be modelled if a stationary linear combination is shown to exist.
%Financial time series are mainly non-stationary.  However, one or more linear
%combinations of non-stationary time series can be stationary even though
%individually they are not. 
When this happens it is said they have a long-run
equilibrium relationship and the variables are {\color{red} cointegrated}.
\

@@

\tikzstyle{every picture}+=[remember picture]
\everymath{\displaystyle}
\begin{frame}
\frametitle{The Vector Error correction}
\tikzstyle{na} = [baseline=-.5ex]

The Vector Error correction model (VECM) is used to model cointegrated time series.

\Large
\begin{equation*}
 \Delta \mathbf{y}_t = \underbrace{
        \tikz[baseline]{
            \node[fill=red!20,ellipse,anchor=base] (t1)
            {$ \Omega\mathbf{y}_{t-1}$};
        }}_\text{Error correction term} +
        \underbrace{
        \tikz[baseline]{
            \node[fill=blue!20,anchor=base] (t2)
            {$\sum_{i=1}^{p-1} \phi_i^* \Delta \mathbf{y}_{t-i} $};
        }}_\text{Autorregresive term} +
        \underbrace{
        \tikz[baseline]{
            \node[fill=green!20,anchor=base] (t3)
            {$c$};
        }}_\text{Constant}
        +
        \underbrace{
        \tikz[baseline]{
            \node[fill=yellow!20,anchor=base] (t3)
            {$\mathbf{\epsilon}_t $};
        }}_\text{i.i.e}
\end{equation*}

\begin{frame}
The Vector Error correction model (VECM) is used to model cointegrated time series but is used only with batch data and rarely used with high frequency data.
\end{frame}

\subsection{Proposal}
\begin{frame}
\frametitle{Proposal}
\begin{block}{}
To use VECM with high frequency data where the best parameters are updated when new data is available and do this is done in a fast way.
\end{block}
\end{frame}

%\section{Background}
%%
%%\begin{frame}
%%\frametitle{Integration }
%%\begin{block}{Integration of order $d$}
%%A time series $\mathbf{y}$ is said to be integrated of order $d$ if after
%%differentiating the variable $d$ times, we get an stationary process, more precisely:
%%\[
%%(1-L)^d \mathbf{y} \sim \text{I(0)} \, ,
%%\]
%%\noindent where I(0) is a stationary time series and $L$ is the lag operator:
%%\[
%%(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
%%\]
%%\end{block}
%%\end{frame}
%%
%%\subsection{Integration example}
%%\begin{frame}
%%\frametitle{Financial Time Series}
%%\begin{columns}
%%\column[t]{5cm}
%%High frequency financial time series are commonly an I(1) process 
%%\begin{center}
%%\includegraphics[width=0.85\textwidth]{img/EURUSD}
%%\end{center}
%%\column[t]{5cm}
%%Since their differences are stationary (I(0) process).
%%\begin{center}
%%\includegraphics[width=0.85\textwidth]{img/DEURUSD}
%%\end{center}
%%\end{columns}
%%\end{frame}
%
%\subsection{Cointegration}
%\begin{frame}
%\frametitle{Cointegration}
%\begin{block}{Cointegration definition}
%Let $\mathbf{y} = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$ be a set of $l$
%time series of order I(1) which are said to be cointegrated if a vector
%$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the
%time series,
%
%\begin{equation*}
% \mathbf{Z}_t:= \beta^\top \mathbf{y} = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
% \text{I(0)}\, .
%\end{equation*}
%\end{block}
%In other words, a set of I(1) variables is said to be cointegrated if
%a linear combination of them exists which is I(0).
%
%\end{frame}
%
%\subsection{Vector Autorregresive models}
%\begin{frame}
%\frametitle{VECM: Vector Error Correction model}
%
%VECM is a special form of a Vector Autorregresive model (VAR) for I(1)
%variables $\mathbf{y}$ that are also cointegrated. Let 
%
%\begin{equation*}
%\label{eq:variables}
%\mathbf{y}_t = 
%\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t}
%\end{bmatrix}^\top \, \qquad t=p+1, \dots N. 
%\end{equation*}
%
%\begin{block}{VAR}
%VAR($p$) model describes the behaviour of a set of $l$
%endogenous variables as a linear combination of their last $p$ values:
%{\color{blue}
%\begin{equation*}
%\label{eq:var}
% \mathbf{y}_t = \phi_1 \mathbf{y}_{t-1}  + \dots +   \phi_p\mathbf{y}_{t-p}
%+ \mathbf{c} + \mathbf{\epsilon}_t \, ,
%\end{equation*}}
%\noindent where  ${\phi_1,\dots,\phi_p}$ are $l \times l$
%matrices of real coefficients,
%$\mathbf{\epsilon}_{p+1},\dots,\mathbf{\epsilon}_N$ are error terms,
%$\mathbf{c}$ is a constant vector and $N$ is the total number of samples.
%\end{block}
%\end{frame}
%
%
%
